---
title: "TidyR and Data Tidying"
author: "Patrick Thane"
date: "2/18/2021"
output:
  html_document:
     toc: true
     toc_depth: 3
     toc_float: true
---

```{r setup, dataframes}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(untidydata)

language_diversity = language_diversity
pre_post = pre_post
```

## Separate
```{r, separate}
pre_post_separate = pre_post %>% 
  separate(col = spec,
           into = c("group", "condition"),
           sep = "_")
```
This is useful for breaking a single column into two columns based upon a predetermined pattern.

## Unite
```{r, unite}
pre_post_unite = pre_post %>% 
  unite(col = "silly", c("id", "spec"), sep = "hello")
```
This is useful for collapsing two columns into a single column.

## Pivot longer
```{r, pivot-longer}
pre_post_pivot = pre_post %>% 
  pivot_longer(cols = c("test1", "test2"),
               names_to = "test",
               values_to = "score")
```
This is useful for placing observations from two columns onto two rows within a single column.

## Pivot wider
```{r, pivot-wider}
language_diversity_wider = language_diversity %>% 
  pivot_wider(
    names_from = Measurement,
    values_from = Value)
```
This is useful for placing observations from two rows into two columns within a single row.